Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-21 paper/main.tex
table

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs on the same problem instances.  As an example, each line in Fig.~\ref{fig:lop_synth} 
shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor? Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
% the fitness of the solutions does not improve monotonically and
new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,
Table~\ref{tab:results} shows\MANUEL{TODO}

%
%    \newcommand{\mcolc}[2]{\multicolumn{#1}{c}{\bf #2}}
%\begin{table}
%  \caption{Results.\label{tab:results}}
%  \resizebox{\textwidth}{!}{%
%\begin{tabular}{*{6}{r}c*{3}{r}}
%  \toprule
%            &                  & \mcolc{4}{Mean fitness (sd)} &                 & \mcolc{2}{Mean runtime}             & \bf     Mean FE\\\cmidrule{3-6}\cmidrule(lr){8-9}
%\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & \bf 95\% CI for the mean difference & \bf UMM  & \bf CEGO                                  & \bf CEGO           \\\midrule
%    LOP     & N-t1d100.01      & 76119.6                      & (915.4)         & 78696.6                             & (811.4)  & (1763.4904574743996, 3390.5095425256004)  & 11436.2 & 260954.5 & 111.8 \\
%    LOP     & N-t1d100.02      & 76827.7                      & (891.0)         & 79686.5                             & (367.9)  & (2194.5410893080507, 3523.058910691955)   & 11353.3 & 262421.6 & 110.7 \\
%    LOP     & N-t1d150.01      & 179508.3                     & (1647.9)        & 187605.5                            & (1906.9) & (6420.259663264915, 9774.140336735109)    & 23395.6 & 261712.4 & 73.9  \\
%    LOP     & N-t1d150.02      & 177592.8                     & (1196.2)        & 184320.9                            & (973.5)  & (5700.461476832616, 7755.738523167395)    & 23361.2 & 263389.9 & 74.1  \\
%    LOP     & N-t1d200.01      & 323513.6                     & (1371.2)        & 335391.1                            & (2289.0) & (10076.07401989604, 13678.92598010396)    & 38476.8 & 263467.5 & 55.6  \\
%    LOP     & N-t1d200.02      & 319675.8                     & (2242.6)        & 333149.5                            & (2497.2) & (11241.971680405779, 15705.428319594244)  & 38268.3 & 265217.5 & 56.5  \\
%    LOP     & N-t2d150.01      & 15297.7                      & (563.5)         & 24682.7                             & (976.6)  & (8622.206776092156, 10147.793223907844)   & 25318.6 & 262324.3 & 74.6  \\
%    LOP     & N-t2d150.02      & 14724.9                      & (667.5)         & 24250.8                             & (1354.9) & (8495.057408450935, 10556.742591549064)   & 25425.6 & 262079.2 & 74.8  \\
%    LOP     & N-t2d200.01      & 33465.2                      & (2001.0)        & 57098.6                             & (4452.8) & (20284.508914101127, 26982.291085898876)  & 40340.9 & 263908.9 & 56.3  \\
%    LOP     & N-t2d200.02      & 35901.8                      & (2077.5)        & 58123.0                             & (2334.9) & (9515.181030361815, 34927.21896963818)    & 40153.1 & 261723.6 & 55.0  \\\midrule
%   PFSP     & rec05            & 1330.7                       & (17.9)          & 1317.9                              & (25.1)   & (-33.424760001214985, 7.8247600012150755) & 506.1   & 139893.0 & 400.0 \\
%   PFSP     & rec13            & 2158.7                       & (29.5)          & 2128.2                              & (25.0)   & (-56.24362499249925, -4.756375007500754)  & 506.8   & 140167.2 & 400.0 \\
%   PFSP     & rec19            & 2405.1                       & (12.6)          & 2398.6                              & (19.8)   & (-22.313869704790818, 9.313869704790818)  & 1078.3  & 295442.7 & 400.0 \\\midrule
%    QAP     & kra32            & 116320.0                     & (4673.5)        & 115907.0                            & (4663.4) & (-4799.295389697205, 3973.295389697205)   & 1194.8  & 344766.2 & 400.0 \\
%    QAP     & nug12            & 660.4                        & (17.3)          & 624.2                               & (26.3)   & (-57.337505259978144, -15.06249474002172) & 180.0   & 53149.3  & 400.0 \\
%    QAP     & nug30            & 7474.0                       & (83.9)          & 7486.8                              & (40.4)   & (-50.84533806724974, 76.4453380672501)    & 1073.5  & 303352.6 & 400.0 \\
%    QAP     & tho30            & 191482.0                     & (4419.7)        & 192416.0                            & (2584.2) & (-2527.0213832396516, 4395.021383239651)  & 1066.5  & 306451.3 & 400.0 \\
%\bottomrule
%\end{tabular}}
%\end{table}

@@

\begin{table}
 \caption{Results.\label{tab:results}}

@@

           &                  & \mcolc{4}{Mean fitness (sd)} &  \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}}         & \mcolc{2}{Mean runtime}             & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}}\\\cmidrule{3-6}\cmidrule(lr){9-10}
\bf Problem & \bf     Instance & \mcolc{2}{UMM}               & \mcolc{2}{CEGO} & & & \bf UMM  & \bf CEGO                                  & \bf            \\\midrule
    LOP &    N-t1d100.01 &       76119.6&(915.4) &        78696.6&(811.4) &       $[$1763.5,&3390.5$]$ &          3.2 &          72.5 &                111.8 \\
    LOP &    N-t1d100.02 &       76827.7&(891.0) &        79686.5&(367.9) &       $[$2194.5,&3523.1$]$ &          3.2 &          72.9 &                110.7 \\
    LOP &    N-t1d150.01 &     179508.3&(1647.9) &      187605.5&(1906.9) &       $[$6420.3,&9774.1$]$ &          6.5 &          72.7 &                 73.9 \\
    LOP &    N-t1d150.02 &     177592.8&(1196.2) &       184320.9&(973.5) &       $[$5700.5,&7755.7$]$ &          6.5 &          73.2 &                 74.1 \\
    LOP &    N-t1d200.01 &     323513.6&(1371.2) &      335391.1&(2289.0) &     $[$10076.1,&13678.9$]$ &         10.7 &          73.2 &                 55.6 \\
    LOP &    N-t1d200.02 &     319675.8&(2242.6) &      333149.5&(2497.2) &     $[$11242.0,&15705.4$]$ &         10.6 &          73.7 &                 56.5 \\
    LOP &    N-t2d150.01 &       15297.7&(563.5) &        24682.7&(976.6) &      $[$8622.2,&10147.8$]$ &          7.0 &          72.9 &                 74.6 \\
    LOP &    N-t2d150.02 &       14724.9&(667.5) &       24250.8&(1354.9) &      $[$8495.1,&10556.7$]$ &          7.1 &          72.8 &                 74.8 \\
    LOP &    N-t2d200.01 &      33465.2&(2001.0) &       57098.6&(4452.8) &     $[$20284.5,&26982.3$]$ &         11.2 &          73.3 &                 56.3 \\
    LOP &    N-t2d200.02 &      35901.8&(2077.5) &       54895.6&(4589.7) &     $[$15539.1,&22448.5$]$ &         11.2 &          73.5 &                 55.6 \\
    LOP &       N-p50-01 &       21027.6&(711.8) &        21431.2&(299.5) &        $[$-128.1,&935.3$]$ &          0.9 &          72.3 &                219.9 \\
    LOP &       N-p50-02 &       21613.6&(596.4) &        22021.1&(275.6) &         $[$-42.5,&857.5$]$ &          0.9 &          72.3 &                217.3 \\
    LOP &       N-atp111 &          594.0&(15.4) &           693.1&(10.3) &          $[$86.7,&111.5$]$ &          4.1 &          72.5 &                 98.7 \\
    LOP &       N-atp134 &          756.3&(21.3) &           880.5&(18.6) &         $[$105.4,&143.0$]$ &          5.7 &          72.8 &                 83.0 \\
    LOP &  N-be75eec\_150 &   1415208.5&(58593.5) &    1666193.1&(72310.8) &   $[$188960.6,&313008.6$]$ &          6.9 &          72.9 &                 74.0 \\
    LOP &  N-be75eec\_250 &  4431271.0&(109756.7) &   5214413.0&(166701.4) &   $[$649040.8,&917243.2$]$ &         15.5 &          73.7 &                 44.9 \\
    LOP &   N-be75np\_150 &   3204878.3&(80704.1) &    3742444.0&(88500.2) &   $[$457944.2,&617187.2$]$ &          6.9 &          72.9 &                 74.2 \\
    LOP &   N-be75np\_250 &  9544754.9&(260523.9) &  10887427.8&(159642.5) & $[$1136635.8,&1548710.0$]$ &         15.5 &          74.0 &                 45.3 \\
   PFSP &      rec05.txt &         1330.7&(17.9) &          1317.9&(25.1) &           $[$-33.4,&7.8$]$ &          0.1 &          38.9 &                400.0 \\
   PFSP &      rec13.txt &         2158.7&(29.5) &          2128.2&(25.0) &          $[$-56.2,&-4.8$]$ &          0.1 &          38.9 &                400.0 \\
   PFSP &      rec19.txt &         2405.1&(12.6) &          2398.6&(19.8) &           $[$-22.3,&9.3$]$ &          0.3 &          82.1 &                400.0 \\
    QAP &      kra32.dat &     116320.0&(4673.5) &      115907.0&(4663.4) &      $[$-4799.3,&3973.3$]$ &          0.3 &          95.8 &                400.0 \\
    QAP &      nug12.dat &          660.4&(17.3) &           624.2&(26.3) &         $[$-57.3,&-15.1$]$ &          0.0 &          14.8 &                400.0 \\
    QAP &      nug30.dat &         7474.0&(83.9) &          7486.8&(40.4) &          $[$-50.8,&76.4$]$ &          0.3 &          84.3 &                400.0 \\
    QAP &      tho30.dat &     191482.0&(4419.7) &      192416.0&(2584.2) &      $[$-2527.0,&4395.0$]$ &          0.3 &          85.1 &                400.0 \\

>>>>

To better understand the behaviour of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an example, each line in Fig.~\ref{fig:lop_synth} 
shows the mean fitness of the solutions evaluated by either uMM or CEGO on one synthetic LOP
instance. The shaded area shows one standard deviation around the
mean. Ideally, each new evaluation will monotonically improve in fitness, since each solution evaluated helps to refine the model and leads to a
better solution being evaluated in the next step.  However, it is often the case that  \EKHINE{yo diría solo que lo esperable es que baje monotonically, sin entrar a valorar un ejemplo concreto}\MANUEL{Así está mejor? Quiero llamar la atención porque normalmente se muestra la evolución de la mejor solución encontrada pero nosotros estamos mostrando el fitness de cada solución evaluada}
% the fitness of the solutions does not improve monotonically and
new solutions
are often worse than their predecessors. Nevertheless, in this example, there
is a clear overall improvement as more evaluations are performed. The minimum
values achieved in the plot up to a particular number of function evaluations
gives an estimation of the mean fitness of the best-so-far solution up to that
point in the run. The algorithms do not return the final solution shown but
rather the best one (minimum fitness) of each run.

@@

Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances). In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
\MANUEL{is the CI for CEGO $-$ UMM or UMM $-$ CEGO? I would prefer}
In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,

@@

\begin{table}[tb]
 \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance. The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO. Mean runtime is measured in seconds. If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 3 days, it was terminated earlier. This was the case for all LOP instances shown.\label{tab:results}}

@@

            &                  & \mcolc{4}{Mean fitness (sd)} & \mcolc{2}{\multirow[b]{2.45}{*}{\shortstack{95\% CI for the\\mean difference}}} & \bf \multirow[b]{2.45}{*}{\shortstack{Mean FE\\CEGO}} & \mcolc{2}{Mean runtime} \\\cmidrule{3-6}\cmidrule(lr){9-10}
\bf Problem & \bf     Instance & \mcolc{2}{CEGO} & \mcolc{2}{UMM}   &                         &            & \bf & \bf CEGO       & \bf UMM                       \\\midrule
    LOP     & N-t1d100.01      & (915.4)         & 76119.6          & 78696.6                 & (811.4)    & $[$1763.5,    & 3390.5$]$     & 111.8& 72.5 & 3.2  \\
    LOP     & N-t1d100.02      & (891.0)         & 76827.7          & 79686.5                 & (367.9)    & $[$2194.5,    & 3523.1$]$     & 110.7& 72.9 & 3.2  \\
    LOP     & N-t1d150.01      & (1647.9)        & 179508.3         & 187605.5                & (1906.9)   & $[$6420.3,    & 9774.1$]$     & 73.9 & 72.7 & 6.5  \\
    LOP     & N-t1d150.02      & (1196.2)        & 177592.8         & 184320.9                & (973.5)    & $[$5700.5,    & 7755.7$]$     & 74.1 & 73.2 & 6.5  \\
    LOP     & N-t1d200.01      & (1371.2)        & 323513.6         & 335391.1                & (2289.0)   & $[$10076.1,   & 13678.9$]$    & 55.6 & 73.2 & 10.7 \\
    LOP     & N-t1d200.02      & (2242.6)        & 319675.8         & 333149.5                & (2497.2)   & $[$11242.0,   & 15705.4$]$    & 56.5 & 73.7 & 10.6 \\
    LOP     & N-t2d150.01      & (563.5)         & 15297.7          & 24682.7                 & (976.6)    & $[$8622.2,    & 10147.8$]$    & 74.6 & 72.9 & 7.0  \\
    LOP     & N-t2d150.02      & (667.5)         & 14724.9          & 24250.8                 & (1354.9)   & $[$8495.1,    & 10556.7$]$    & 74.8 & 72.8 & 7.1  \\
    LOP     & N-t2d200.01      & (2001.0)        & 33465.2          & 57098.6                 & (4452.8)   & $[$20284.5,   & 26982.3$]$    & 56.3 & 73.3 & 11.2 \\
    LOP     & N-t2d200.02      & (2077.5)        & 35901.8          & 54895.6                 & (4589.7)   & $[$15539.1,   & 22448.5$]$    & 55.6 & 73.5 & 11.2 \\
    LOP     & N-p50-01         & (711.8)         & 21027.6          & 21431.2                 & (299.5)    & $[$-128.1,    & 935.3$]$      & 219.9& 72.3 & 0.9  \\
    LOP     & N-p50-02         & (596.4)         & 21613.6          & 22021.1                 & (275.6)    & $[$-42.5,     & 857.5$]$      & 217.3& 72.3 & 0.9  \\
    LOP     & N-atp111         & (15.4)          & 594.0            & 693.1                   & (10.3)     & $[$86.7,      & 111.5$]$      & 98.7 & 72.5 & 4.1  \\
    LOP     & N-atp134         & (21.3)          & 756.3            & 880.5                   & (18.6)     & $[$105.4,     & 143.0$]$      & 83.0 & 72.8 & 5.7  \\
    LOP     & N-be75eec\_150   & (58593.5)       & 1415208.5        & 1666193.1               & (72310.8)  & $[$188960.6,  & 313008.6$]$   & 74.0 & 72.9 & 6.9  \\
    LOP     & N-be75eec\_250   & (109756.7)      & 4431271.0        & 5214413.0               & (166701.4) & $[$649040.8,  & 917243.2$]$   & 44.9 & 73.7 & 15.5 \\
    LOP     & N-be75np\_150    & (80704.1)       & 3204878.3        & 3742444.0               & (88500.2)  & $[$457944.2,  & 617187.2$]$   & 74.2 & 72.9 & 6.9  \\
    LOP     & N-be75np\_250    & (260523.9)      & 9544754.9        & 10887427.8              & (159642.5) & $[$1136635.8, & 1548710.0$]$  & 45.3 & 74.0 & 15.5 \\
   PFSP     & rec05.txt        & (17.9)          & 1330.7           & 1317.9                  & (25.1)     & $[$-33.4,     & 7.8$]$        & 400.0& 38.9 & 0.1  \\
   PFSP     & rec13.txt        & (29.5)          & 2158.7           & 2128.2                  & (25.0)     & $[$-56.2,     & -4.8$]$       & 400.0& 38.9 & 0.1  \\
   PFSP     & rec19.txt        & (12.6)          & 2405.1           & 2398.6                  & (19.8)     & $[$-22.3,     & 9.3$]$        & 400.0& 82.1 & 0.3  \\
    QAP     & kra32.dat        & (4673.5)        & 116320.0         & 115907.0                & (4663.4)   & $[$-4799.3,   & 3973.3$]$     & 400.0& 95.8 & 0.3  \\
    QAP     & nug12.dat        & (17.3)          & 660.4            & 624.2                   & (26.3)     & $[$-57.3,     & -15.1$]$      & 400.0& 14.8 & 0.0  \\
    QAP     & nug30.dat        & (83.9)          & 7474.0           & 7486.8                  & (40.4)     & $[$-50.8,     & 76.4$]$       & 400.0& 84.3 & 0.3  \\
    QAP     & tho30.dat        & (4419.7)        & 191482.0         & 192416.0                & (2584.2)   & $[$-2527.0,   & 4395.0$]$     & 400.0& 85.1 & 0.3  \\
